### 2023-06-21 abstract.tex
section abstract

\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
We consider six types of resilience (one of them is the home-state property) defined by the two subsets Bad and Safe and by the existence of a bound on the length of runs starting from Bad and reaching Safe. We study the decidability of each type of resilience for two kinds of models: WSTS and VASS. We first show that most of all resilience problems are undecidable for WSTS with strong compatibility and with effective pred-basis. Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility. Moreover, some resilience properties are decidable for two other classes of WSTS : (1) for WSTS with effective 
$\uparrow$ $\post^*$ basis and (2) for ideal-effective WSTS with downward and upward compatibilities. Finally, we study the resilience for VASS where most of the problems are shown decidable.

>>>>

\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
We consider six types of resilience (one of them is the home-space property) defined by Bad is an upward/downward closed set or it is a subset of the reachability set from an initial state, Safe is upward-closed ($\Safe=\uparrow \Safe$) or Safe is downward-closed ($\Safe=\downarrow \Safe$) (Bad is the complementary of Safe) and by the existence (or not) of a bound $k$ on the length of runs starting from Bad and reaching Safe. We study the decidability of each type of resilience for two kinds of models: WSTS and VASS. We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective pred-basis). Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$. Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective 
$\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$. Finally, we study the resilience for VASS with semi-linear subsets Bad and Safe and for extensions and variations of VASS; most of the resilience properties are shown decidable.

### 2023-06-21 resilience-safe-up.tex
section 3
section abstract

We start with the case $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$. 
It can be viewed as a generalizeation
 of coverability, as
it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.

@@

Without the completion-post-effective hypothesis nor the pred-basis hypothesis, coverability 
becomes undecidable. For instance the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}. 
Coverability reduce
to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$,
as $s \to^* \uparrow t$ 
reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
Hence, without the completion-post-effective hypothesis, 
{\sc Resilience} itself is undecidable.

Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using
$\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 
the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.

>>>>

We start with the case $\Safe=\uparrow \Safe$ hence $\Bad=\downarrow \Bad$. 
It can be viewed as a generalizeation
 of coverability, as
it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.

@@

Let us remark that coverability reduce
to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$,
as $s \to^* \uparrow t$ 
reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
%
Let us recall that the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}, hence the strongly increasing well-structured nets are not completion-post-effective hypothesis nor have the pred-basis hypothesis.
%
Without the completion-post-effective hypothesis nor the pred-basis hypothesis \alain{où est-ce prouvé ?}, coverability 
becomes undecidable. For instance

Hence, without the completion-post-effective hypothesis, 
{\sc Resilience} itself is undecidable.
\alain{c'est confus}
Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using
$\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 
the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.

### 2023-06-21 resilience.tex
up
section abstract

@@

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$. In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
Remark that in Example~\ref{Example}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@

{\bf Synthesis of the main decidability results}\label{synthesis}
\begin{center}
\begin{tabular}{ | l | c | c | c | r |}
\hline   \Safe~\Bad %& $\uparrow$~ $\uparrow$~ 
		& $\uparrow$~ $\downarrow$~ 
		 & $\downarrow$~ $\uparrow$~ 
		 %& $\downarrow$~ $\downarrow$~
 \\ \hline
   RP %& Decidable (Thm~\ref{up-up}) 
   	& Decidable (Thm~\ref{down-up})  
   		 & Decidable (Thm~\ref{up-down}) 
   	%	 & Undecidable (Thm~\ref{down-down})
    \\ \hline
   BRP %& Decidable (Corollary~\ref{B-up-up}) 
   &  Decidable (Thm~\ref{down-up}) 
   		 & ?? 
   	%	 & ??
    \\ \hline
      kRP %& Decidable (Thm~\ref{k-up-up}) 
      & Decidable (Thm~\ref{down-up}) 
      		& ?? 
      	%	& ??
       \\ \hline
 \end{tabular}
\end{center}

@@

\begin{proof}
Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.

It is also possible to define $(s_0,X)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} X$.

>>>>

\mathieu{Rappeler l'example  les Safe Bad considérés et dire ce que ça veut dire en termes de chemins etc.}

Recall the {\bf Flight Controller System} example, the safe state corresponds to the aircrafts having both landed, i.e.
$\Safe = \{(q_{A,1}, q_{M,5}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4\}$,
and $\Bad$ is the complement of $\Safe$. 
% $\Bad = \{(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{v}) \mid \textbf{v} \in \N^4\}$.
%%  Then the resilience problem asks
% wheher

@@

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$. In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
Remark that in Example~\ref{Example} ({(\bf Flight Controller System)}), $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@

{\bf Synthesis of the main decidability results}\label{synthesis}
\begin{center}
\begin{tabular}{ | l | c | c | c | r |}
\hline   \ %& $\uparrow$~ $\uparrow$~ 
		& $\Safe = {\uparrow \Safe}$~ 
		 & $\Safe = {\downarrow \Safe}$~ 
		 %& $\downarrow$~ $\downarrow$~
 \\ \hline
   RP %& Decidable (Thm~\ref{up-up}) 
   	& Decidable (Thm~\ref{down-up})  
   		 & Decidable (Thm~\ref{up-down}) 
   	%	 & Undecidable (Thm~\ref{down-down})
    \\ \hline
   BRP %& Decidable (Corollary~\ref{B-up-up}) 
   &  Decidable (Thm~\ref{down-up}) 
   		 & ?? 
   	%	 & ??
    \\ \hline
      kRP %& Decidable (Thm~\ref{k-up-up}) 
      & Decidable (Thm~\ref{down-up}) 
      		& ?? 
      	%	& ??
       \\ \hline
 \end{tabular}
\end{center}

@@

\begin{proof}
Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
%		\end{proof}
\alain{impossible de mettre end{proof} ???...et la preuve ne se finit pas clairement}
%
It is also possible to define $(s_0,X)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} X$.
